---
title: "R_SQL_performance_cheatsheet"
output: html_notebook
---
```{r}
library(stats)
library(ISLR)
library(MASS)

tit_train = read.csv("C:/Users/sin17h/Documents/datasc/kaggle/Titanic_classify/data/train.csv")
tit_test = read.csv("C:/Users/sin17h/Documents/datasc/kaggle/Titanic_classify/data/test.csv")

tit_train$Title <- gsub('(.*, )|(\\..*)', '', tit_train$Name)

```

Ingest data into SQL server
```{r}

library(DBI)
library(odbc)
sort(unique(odbcListDrivers()[[1]]))

con <- dbConnect(odbc::odbc(), 
                 Driver = "SQL Server", 
                 Server = "localhost\\sql2019", 
                 Database = "Titanic", 
                 Trusted_Connection = "True")

dbWriteTable(con, "tit_train", tit_train)
```

```{r}

table(tit_train$Title)
#equivalent SQL
# select Title, count(*) as cnt
# from tit_train
# group by Title

table(tit_train$Sex, tit_train$Title)
# select Title
#       ,count(case when Sex='female' then 1 else 0 end) as 'female'
#       ,count(case when Sex='male' then 1 else 0 end) as 'female'
# from tit_train
# group by Title

tapply(tit_train$Sex, tit_train$Title, length) 
# select Title, count(Sex) as cnt
# from tit_train
# group by Title

by(tit_train,tit_train$Title,length)

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
