---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(tidylog)
library(heatmaply)
getwd()
```


Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

Read the titanic data set. And summarize it.
```{r}
tit_train = read.csv("data/train.csv")
str(tit_train)
summary(tit_train)
```
Summary tells us that:
1. Passenger id has correlation with survivability or not. 
   cor(tit_train$PassengerId,tit_train$Survived) is -0.005006661, so perhaps not

2. nulls/NAs in Age column that we will deal later. how to deal: remove the rows or mean or 0 or use the parent-child info from the dataset to get the mean of that category.
   length(tit_train$Age[is.na(tit_train$Age)]) is 177. removing the rows and using cor gave a value of -0.07722109 which is close to 0.
   
```{r}
# heatmaply_cor(
#   cor(tit_train),
#   xlab = "Features",
#   ylab = "Features",
#   k_col = 2,
#   k_row = 2
# )

cor(tit_train$PassengerId,tit_train$Survived)


par(mfrow=c(3,3))
length(tit_train$Age[is.na(tit_train$Age)])
cor(tit_train$Age,tit_train$Survived) #this wont work because of null values. so temporarily, use subset to sniff the relationship
sb_tit = subset(tit_train,subset = !is.na(tit_train$Age))#this reduces the dataset to 714 rows.
plot(sb_tit$Survived, sb_tit$Age)
boxplot(sb_tit$Age~sb_tit$Survived)
cor(sb_tit$Age,sb_tit$Survived) #-0.07722109. So perhaps its suggest that there is not correlation?

plot(tit_train$Age)
hist(tit_train$Age)
boxplot(sb_tit$Age)
plot(tit_train$Survived, tit_train$Age)
boxplot(tit_train$Age~tit_train$Survived)
```
3. Name has a unique value for each observation. Can be disregarded as it should not have an affect on survivability. Check by making a plot. But first reduce the textual value to a number by coding it. options are: discard, hash to a numeric value or do feature engineering to extract create family linkages (using last name) 
```{r}
#cor(tit_train$Name,tit_train$Survived)# name has to be encoded
head(tit_train$Name)
tit_train$Title <- gsub('(.*, )|(\\..*)', '', tit_train$Name)
unique(Title)

table(Title)#group by and count/length
table(tit_train$Sex,tit_train$Title) #table is equivalent to groupby with implicit count/length using a case statement
tapply(tit_train$Sex,tit_train$Title,length) # this is equivalent to table(Title). The equivalent group is has implicit count/length without case statement
sort(tapply(tit_train$Sex,tit_train$Title,length))

# Titles with very low cell counts to be combined to "rare" level
rare_title = c('Dona', 'Lady', 'the Countess','Capt', 'Col', 'Don', 
                'Dr', 'Major', 'Rev', 'Sir', 'Jonkheer')
# Also reassign mlle, ms, and mme accordingly
tit_train$Title[tit_train$Title == 'Mlle']        <- 'Miss' #update table where condition
tit_train$Titletit_train[tit_train$Title == 'Ms']          <- 'Miss' #update table where condition
tit_train$Title[tit_train$Title == 'Mme']         <- 'Mrs'  #update table where condition
tit_train$Title[tit_train$Title %in% rare_title]  <- 'Rare Title' #update table where condition
table(tit_train$Sex,tit_train$Title) 

# Finally, grab surname from passenger name
strsplit('Braund, Mr. Owen Harris', split = '[,.]')[[1]][1] #returns 'Braund'
#this does not work as strsplit only works on strings whereas tit_train$Name is a factor. So first typecast it to character vector(aka string)
#tit_train$Surname <- sapply(tit_train$Name, function(x) strsplit(x, split = '[,.]')[[1]][1])
str(tit_train$Name); class(tit_train$Name)
tit_train$Surname <- sapply(as.character(tit_train$Name), function(x) strsplit(x, split = '[,.]')[[1]][1])

length(tit_train$Surname) #891
length(unique(tit_train$Surname)) #667
cat(paste('We have <b>', nlevels(factor(tit_train$Surname)), '</b> unique surnames. I would be interested to infer ethnicity based on surname --- another time.'))

#a = subset(tit_train$Name, grepl("the Countess",tit_train$Name))
#length(a)
```

4. Let create a family size column
```{r}
tit_train$Fsize = tit_train$SibSp + tit_train$Parch + 1
tit_train$Family = paste(tit_train$Surname, tit_train$Fsize, sep = "_")

par(mfrow=c(2,2))
plot(tit_train$Survived, tit_train$Fsize)
boxplot(tit_train$Fsize~tit_train$Survived)
hist(tit_train$Fsize[tit_train$Survived==1], breaks = 11)
hist(tit_train$Fsize[tit_train$Survived==0], breaks = 11)
summary(tit_train$Fsize[tit_train$Survived==1])
summary(tit_train$Fsize[tit_train$Survived==0])

par(mfrow=c(1,1))
# Use ggplot2 to visualize the relationship between family size & survival. I could not produce a simialr plot with hist 
# which can handle factor/levels
ggplot(tit_train[1:891,], aes(x = Fsize, fill = factor(Survived))) +
  geom_bar(stat='count', position='dodge') +
  scale_x_continuous(breaks=c(1:11)) +
  labs(x = 'Family Size') 
  #+ theme_few()

# Discretize family size
tit_train$FsizeD[tit_train$Fsize == 1] <- 'singleton'
tit_train$FsizeD[tit_train$Fsize < 5 & tit_train$Fsize > 1] <- 'small'
tit_train$FsizeD[tit_train$Fsize > 4] <- 'large'

# Show family size by survival using a mosaic plot. remeber table groups by and use count(or length agregate func). Sowhat we see 
#here is the count of families with a case statement on family size
mosaicplot(table(tit_train$FsizeD, tit_train$Survived), main='Family Size by Survival', shade=TRUE)
# heatmaply(tit_train)
```

5. Now take a look at Cabin. It has a lot of empty string as value but they don't show up in summary.
```{r}
str(tit_train$Cabin)
head(sort(tit_train$Cabin))
tail(sort(tit_train$Cabin))

training_tit = read.csv("data/test.csv")
fix(training_tit)
head(sort(training_tit$Cabin))
tail(sort(training_tit$Cabin))
summary(training_tit)
```



But now lets create a hist plot for each feature and then also facet chart

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.
Before we 
```{r}
hist(tit_train$Survived)
pairs(tit_train)
```


When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
