---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(tidylog)
library(heatmaply)
library(ggplot2)
library(GGally)
library('ggthemes') # visualization
library('scales') # visualization
library('dplyr') # data manipulation
library('mice') # imputation
library('randomForest') # classification algorithm
getwd()
```


Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

Read the titanic data set. And summarize it.
```{r}
tit_train = read.csv("data/train.csv")
tit_test = read.csv("data/test.csv")
str(tit_train)
str(tit_test)
summary(tit_train)
md.pattern(tit_train)#this as well does not give insight about issing values in df.

full = bind_rows(tit_train,tit_test)
str(full)
summary(full)
#fix(full)
```
Summary tells us that:
1. Passenger id has correlation with survivability or not. 
   cor(full$PassengerId,full$Survived) is -0.005006661, so perhaps not

2. nulls/NAs in Age column that we will deal later. how to deal: remove the rows or mean(/median/mode) or 0 or use the parent-child info from the dataset to get the mean of that category.
   length(full$Age[is.na(full$Age)]) is 177. removing the rows and using cor gave a value of -0.07722109 which is close to 0.
   
```{r}
# heatmaply_cor(
#   cor(full),
#   xlab = "Features",
#   ylab = "Features",
#   k_col = 2,
#   k_row = 2
# )

cor(full$PassengerId,full$Survived)


par(mfrow=c(3,3))
length(full$Age[is.na(full$Age)])
cor(full$Age,full$Survived) #this wont work because of null values. so temporarily, use subset to sniff the relationship
sb_tit = subset(full,subset = !is.na(full$Age))#this reduces the dataset to 714 rows.
plot(sb_tit$Survived, sb_tit$Age)
boxplot(sb_tit$Age~sb_tit$Survived)
cor(sb_tit$Age,sb_tit$Survived) #-0.07722109. So perhaps its suggest that there is not correlation?

plot(full$Age)
hist(full$Age)
boxplot(sb_tit$Age)
plot(full$Survived, full$Age)
boxplot(full$Age~full$Survived)
```
3. Name has a unique value for each observation. Can be disregarded as it should not have an affect on survivability. Check by making a plot. But first reduce the textual value to a number by coding it. options are: discard, hash to a numeric value or do feature engineering to extract create family linkages (using last name) 
```{r}
#cor(full$Name,full$Survived)# name has to be encoded
head(full$Name)
full$Title <- gsub('(.*, )|(\\..*)', '', full$Name)
unique(full$Title)

table(full$Title)#group by and count/length
table(full$Sex,full$Title) #table is equivalent to groupby with implicit count/length using a case statement
tapply(full$Sex,full$Title,length) # this is equivalent to table(Title). The equivalent group is has implicit count/length without case statement
sort(tapply(full$Sex,full$Title,length))

# Titles with very low cell counts to be combined to "rare" level
rare_title = c('Dona', 'Lady', 'the Countess','Capt', 'Col', 'Don', 
                'Dr', 'Major', 'Rev', 'Sir', 'Jonkheer')
# Also reassign mlle, ms, and mme accordingly
full$Title[full$Title == 'Mlle']        <- 'Miss' #update table where condition
full$Title[full$Title == 'Ms']          <- 'Miss' #update table where condition
full$Title[full$Title == 'Mme']         <- 'Mrs'  #update table where condition
full$Title[full$Title %in% rare_title]  <- 'Rare Title' #update table where condition
table(full$Sex,full$Title) 

# Finally, grab surname from passenger name
strsplit('Braund, Mr. Owen Harris', split = '[,.]')[[1]][1] #returns 'Braund'
#this does not work as strsplit only works on strings whereas full$Name is a factor. So first typecast it to character vector(aka string)
#full$Surname <- sapply(full$Name, function(x) strsplit(x, split = '[,.]')[[1]][1])
str(full$Name); class(full$Name)
full$Surname <- sapply(as.character(full$Name), function(x) strsplit(x, split = '[,.]')[[1]][1])

length(full$Surname) #891
length(unique(full$Surname)) #667
cat(paste('We have <b>', nlevels(factor(full$Surname)), '</b> unique surnames. I would be interested to infer ethnicity based on surname --- another time.'))

#a = subset(full$Name, grepl("the Countess",full$Name))
#length(a)
```

4. Let create a family size column
```{r}
full$Fsize = full$SibSp + full$Parch + 1
full$Family = paste(full$Surname, full$Fsize, sep = "_")

par(mfrow=c(2,2))
plot(full$Survived, full$Fsize)
boxplot(full$Fsize~full$Survived)
hist(full$Fsize[full$Survived==1], breaks = 11)
hist(full$Fsize[full$Survived==0], breaks = 11)
summary(full$Fsize[full$Survived==1])
summary(full$Fsize[full$Survived==0])

par(mfrow=c(1,1))
# Use ggplot2 to visualize the relationship between family size & survival. I could not produce a simialr plot with hist 
# which can handle factor/levels
ggplot(full[1:891,], aes(x = Fsize, fill = factor(Survived))) +
  geom_bar(stat='count', position='dodge') +
  scale_x_continuous(breaks=c(1:11)) +
  labs(x = 'Family Size') 
  #+ theme_few()

# Discretize family size
full$FsizeD[full$Fsize == 1] <- 'singleton'
full$FsizeD[full$Fsize < 5 & full$Fsize > 1] <- 'small'
full$FsizeD[full$Fsize > 4] <- 'large'

# Show family size by survival using a mosaic plot. remeber table groups by and use count(or length agregate func). Sowhat we see 
#here is the count of families with a case statement on family size
mosaicplot(table(full$FsizeD, full$Survived), main='Family Size by Survival', shade=TRUE)
# heatmaply(full)
```

5. Now take a look at Cabin. It has a lot of empty string as value but they don't show up in summary. In python, df.info() shows us the gives us a better insight as far as not-null data is concerned
```{r}
str(full$Cabin)
head(sort(full$Cabin))
tail(sort(full$Cabin))
summary(full$Cabin)#does not show empty strings.
length(full$Cabin[full$Cabin == ''])#out of 891, we have 687 empty strings. So how we look for this anomaly?

# Create a Deck variable. Get passenger deck A - F:
full$Deck<-factor(sapply(as.character(full$Cabin), function(x) strsplit(x, NULL)[[1]][1]))
str(full$Deck)

```
6. Lets deal with missing data. The que`stion remains: i have dont know a high level func that will point out the columns that have missing values in R dataframe.

```{r}
par(mfrow=c(3,3))

full$Embarked  = as.factor(full$Embarked)
summary(full$Embarked)
plot(full$Embarked)
which(full$Embarked=='')#we can build a custom func that loops around this to give us the desired result
full$Embarked[63]
full[c(62, 830), 'Embarked']
cat(paste('We will infer their values for **embarkment** based on present data that we can imagine may be relevant: **passenger class** and **fare**. We see that they paid<b> $', full[c(62, 830), 'Fare'][1], '</b>and<b> $', full[c(62, 830), 'Fare'][2], '</b>respectively and their classes are<b>', full[c(62, 830), 'Pclass'][1], '</b>and<b>', full[c(62, 830), 'Pclass'][2], '</b>. So from where did they embark?'))

full[c(62, 830), 'Fare'][[1]][1]
full$Fare[c(62,830)][]
full[62,]

#lets see if we can infer the embark values from other rows
pairs(~Fare + Pclass + Embarked ,full)
plot(full$Embarked, full$Fare) 

plot(full$Pclass, full$Fare)
plot(as.factor(full$Pclass), full$Fare)
plot(full$Pclass, full$Embarked)

#what we can get using pair above is fine. But striaghtforwards way to plot on a single plot 3 with ggplot is:
# Use ggplot2 to visualize embarkment, passenger class, & median fare
#so they embarked from C
ggplot(full, aes(x = Embarked, y = Fare, fill = factor(Pclass))) +
  geom_boxplot() +
  geom_hline(aes(yintercept=80), 
    colour='red', linetype='dashed', lwd=2) +
  scale_y_continuous(labels=dollar_format()) +
  theme_few()

full$Embarked[c(62, 830)] = 'C'
```

7. Fare also has NA value. replace the na value with median of sample with same class as this passenger as well as same embarkment point
```{r}
str(full$Fare)
summary(full$Fare)
which(is.na(full$Fare))
full[1044,]
sb = subset(full,full$Embarked=='S' & full$Pclass == 3)
summary(sb$Fare)["Median"]
plot(sb$Fare)
hist(sb$Fare)#hist shows the distribution 
ggplot(full[full$Pclass == '3' & full$Embarked == 'S', ], 
  aes(x = Fare)) +
  geom_density(fill = '#99d6ff', alpha=0.4) + 
  geom_vline(aes(xintercept=median(Fare, na.rm=TRUE)),
    colour='red', linetype='dashed', lwd=1) +
  scale_x_continuous(labels=dollar_format()) +
  theme_few()

 full$Fare[1044] = summary(sb$Fare)["Median"]
full$Fare[1044] = median(summary(sb$Fare))
                         
```

8. Now lets fix age variable which has
```{r}

# Show number of missing Age values
sum(is.na(full$Age))
#length(is.na(full$Age)) # length wont work here because we haven't filtered the rows 
length(full$Age[is.na(full$Age)]) # but it would work if we filter the rows

plot(full$Age)
hist(full$Age)
summary(full$Age)

# Make variables factors into factors
factor_vars <- c('PassengerId','Pclass','Sex','Embarked',
                 'Title','Surname','Family','FsizeD')

summary(full[factor_vars])
full[factor_vars] <- lapply(full[factor_vars], function(x) as.factor(x))
summary(full[factor_vars])

# Set a random seed
set.seed(129)

# Perform mice imputation, excluding certain less-than-useful variables:
mice_mod <- mice(full[, !names(full) %in% c('PassengerId','Name','Ticket','Cabin','Family','Surname','Survived')], method='rf') 

# Save the complete output 
mice_output <- complete(mice_mod)

# Plot age distributions
par(mfrow=c(1,2))
hist(full$Age, freq=F, main='Age: Original Data', 
  col='darkgreen', ylim=c(0,0.04))
hist(mice_output$Age, freq=F, main='Age: MICE Output', 
  col='lightgreen', ylim=c(0,0.04))

# Replace Age variable from the mice model.
full$Age <- mice_output$Age

# Show new number of missing Age values
sum(is.na(full$Age))

```

9. let's look at relationship betwee age,survival and sex. The problem is we can't can't see relationship between 3 variables without ggplot.
```{r}
pairs(~full$Age+full$Sex+full$Survived)
plot(full$Survived,full$Age)
# First we'll look at the relationship between age & survival
ggplot(full[1:891,], aes(Age, fill = factor(Survived))) + 
  geom_histogram() + 
  # I include Sex since we know (a priori) it's a significant predictor
  facet_grid(.~Sex) + 
  theme_few()

# Create the column child, and indicate whether child or adult
full$Child[full$Age < 18] <- 'Child'
full$Child[full$Age >= 18] <- 'Adult'

# Show counts
table(full$Child, full$Survived)

# Adding Mother variable
full$Mother <- 'Not Mother'
full$Mother[full$Sex == 'female' & full$Parch > 0 & full$Age > 18 & full$Title != 'Miss'] <- 'Mother'

# Show counts
table(full$Mother, full$Survived)

# Finish by factorizing our two new factor variables
full$Child  <- factor(full$Child)
full$Mother <- factor(full$Mother)
```

10. lets look at missing values. I do not think this again gives an idea of blank strings. But even then, it should be used with summary to look at NA patterns in the dataframe.
```{r}
md.pattern(full)

```

11. do the prediction now.
```{r}

# Split the data back into a train set and a test set
train <- full[1:891,]
test <- full[892:1309,]

# Set a random seed
set.seed(754)

# Build the model (note: not all possible variables are used)
rf_model <- randomForest(factor(Survived) ~ Pclass + Sex + Age + SibSp + Parch + 
                                            Fare + Embarked + Title + 
                                            FsizeD + Child + Mother,
                                            data = train)

# Show model error
plot(rf_model, ylim=c(0,0.36)) + legend('topright', colnames(rf_model$err.rate), col=1:3, fill=1:3)

```

12. variable importance:
```{r}
# Get importance
importance    <- importance(rf_model)
varImportance <- data.frame(Variables = row.names(importance), 
                            Importance = round(importance[ ,'MeanDecreaseGini'],2))

# Create a rank variable based on importance
rankImportance <- varImportance %>%
  mutate(Rank = paste0('#',dense_rank(desc(Importance))))

# Use ggplot2 to visualize the relative importance of variables
ggplot(rankImportance, aes(x = reorder(Variables, Importance), 
    y = Importance, fill = Importance)) +
  geom_bar(stat='identity') + 
  geom_text(aes(x = Variables, y = 0.5, label = Rank),
    hjust=0, vjust=0.55, size = 4, colour = 'red') +
  labs(x = 'Variables') +
  coord_flip() + 
  theme_few()
```
13. Prediction :
```{r}
# Predict using the test set
prediction <- predict(rf_model, test)

# Save the solution to a dataframe with two columns: PassengerId and Survived (prediction)
solution <- data.frame(PassengerID = test$PassengerId, Survived = prediction)

# Write the solution to file
write.csv(solution, file = 'rf_mod_Solution.csv', row.names = F)
```


But now lets create a hist plot for each feature and then also facet chart

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.
Before we 
```{r}
hist(full$Survived)
pairs(full)
```


When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.


