---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```


Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

Read the titanic data set. And summarize it.
```{r}
tit_train = read.csv("C:/Users/sin17h/Documents/datasc/kaggle/Titanic_classify/train.csv")
str(tit_train)
summary(tit_train)
```
Summary tells us that:
1. Passenger id has correlation with survivability or not. 
   cor(tit_train$PassengerId,tit_train$Survived) is -0.005006661, so perhaps not

2. nulls/NAs in Age column that we will deal later. how to deal: remove the rows or mean or 0 or use the parent-child info from the dataset to get the mean of that category.
   length(tit_train$Age[is.na(tit_train$Age)]) is 177. removing the rows and using cor gave a value of -0.07722109 which is close to 0.
   
```{r}
par(mfrow=c(3,3))
cor(tit_train$PassengerId,tit_train$Survived)

length(tit_train$Age[is.na(tit_train$Age)])
cor(tit_train$Age,tit_train$Survived) #this wont work because of null values. so temporarily, use subset to sniff the relationship
sb_tit = subset(tit_train,subset = !is.na(tit_train$Age))#this reduces the dataset to 714 rows.
plot(sb_tit$Survived, sb_tit$Age)
boxplot(sb_tit$Age~sb_tit$Survived)
cor(sb_tit$Age,sb_tit$Survived) #-0.07722109. So perhaps its suggest that there is not correlation?

plot(tit_train$Age)
hist(tit_train$Age)
boxplot(sb_tit$Age)
plot(tit_train$Survived, tit_train$Age)
boxplot(tit_train$Age~tit_train$Survived)
```
3. Name has a unique value for each observation. Can be disregarded as it should not have an affect on survivability. Check by making a plot. But first reduce the textual value to a number by coding it. options are: discard, hash to a numeric value or do feature engineering to extract create family linkages (using last name) 
```{r}
#cor(tit_train$Name,tit_train$Survived)# name has to be encoded
head(tit_train$Name)
tit_train$Title <- gsub('(.*, )|(\\..*)', '', tit_train$Name)
unique(Title)

table(Title)
table(tit_train$Sex,tit_train$Title)

#a = subset(tit_train$Name, grepl("the Countess",tit_train$Name))
#length(a)
```


But now lets create a hist plot for each feature and then also facet chart

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.
Before we 
```{r}
hist(tit_train$Survived)
pairs(tit_train)
```


When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
